### 2021-10-09 slides_modelers-meeting.Rmd
Add results and summary

---
title: "Sampling in Decisions from Experience"
subtitle: https://github.com/linushof/sampling-in-dfe
author: "Linus, Veronika & Thorsten"
date: "October 11, 2021"
output:
  beamer_presentation: default
bibliography: references_sampling-in-dfe.bib
csl: apa.csl
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
pacman::p_load(tidyverse, 
               scico,
               ggpubr)
```

## Project Description

### Starting Ideas

1. Description-experience gap [cf. @wulffMetaanalyticReviewTwo2018]

2. Empirical samples that carry over to cognitive processes

### Core Question

- May samples be generated in such a way that they produce systematic choice patterns?

## Prospects and Sampling

- Standard paradigm: Choice between 2 or multiple prospects

- *Prospects*: Probability spaces $(\Omega, \mathcal{F}, P)$, where 
$$
\Omega = \{\omega_1, ..., \omega_n\}
$$
and 
$$
P = \{p(\omega_1), ..., p(\omega_n)\}
$$

- *Single sample*: Realization of a random variable:
$$
X(\omega \in \Omega) = x 
$$

- *Sequential sampling*: Generating a sequence of single samples in discrete time: 
$$
X_i = X_1, ..., X_n 
$$

## Difference between DfD and DfE

- Decisions from Description (DfD): Function of symbolical descriptions of prospects:  
$$
D:= f((\Omega_1, \mathcal{F}_1, P_1), ..., (\Omega_k, \mathcal{F}_k, P_k))
$$

- Decisions from Experience (DfE): Function of sample sequences originating from prospects:
$$
D := f(X_{i1}, ..., X_{ik}) 
$$

- $f$ informed by psychological theory and empirical protocols

## Sampling and Decision Strategies

- Hills and Hertwig [-@hillsInformationSearchDecisions2010]: Assumption of a systematic link between sampling and decision strategies

- Sampling strategy: Succession of single samples in a sequence generated from multiple prospects

  + Piecewise: Single samples from **different** prospects 
  + Comprehensive: Single samples from **same** prospect

```{r, echo=FALSE, out.height="40%", out.width="40%", fig.align='center'}
knitr::include_graphics("hills-hertwig_2010_figure-1.png")
```

## Modeling the Assumptions of Hills & Hertwig (2010)

- Consider two prospects $(\Omega_X, \mathcal{F}_X, P_X)$ and $(\Omega_Y, \mathcal{F}_Y, P_Y)$ and the ground model for DfE
$$
D := f(X_i, Y_i) 
$$

- $f$ maps a comparison of $\overline{X}$ and $\overline{Y}$ on the set $\{0, 1\}$, with 0 (1) indicating a lost (won) comparison:
$$
D := f(\overline{X} - \overline{Y}) = 
  \begin{cases}
    1 \quad \text{for} \quad \overline{X} - \overline{Y} > 0 \\
    0 \quad \text{else}. 
  \end{cases} 
$$

- A series of ordinal comparisons $D_i = D_1, ..., D_n$ is a Bernoulli process where the number of won comparisons follows the binomial distribution
$$
D \sim \mathcal{B}\left( p \left(\overline{X} - \overline{Y}> 0\right), n\right)
$$

## Predicting Choices from the Stochastic Model

- $n$ and $p$ of the binomial distribution are influenced by the sampling and decision strategies:

### Comprehensive Sampling and Summary Decisions

- $n = 1$  
- $p$: For $\lim_{N_X, N_Y  \to \infty}$, $p (\overline{X} - \overline{Y}> 0) \approx p (E(X) - E(Y))$ 
- Choices should become *"normative"*, i.e., linear weighting and value function

### Piecewise Sampling and Round-wise Decisions

- $n \geq 1$ and $N_X, N_Y = 1$
- $p$: For $N_X, N_Y = 1$, $p (\overline{X} - \overline{Y} > 0) = p(\omega \in \Omega_X > \omega \in \Omega_Y)$
- Choices depend on probabilities in the mid-range, i.e., S-shaped weighting function and concave value function

## Method

- 60 choices between a 2-outcome prospect and a safe prospect
- 100 synthetic agents
- Parameterization of the sampling and decision process
  + Probability of switching between prospects
  + Number of comparisons
- Description of simulated DfE in Cumulative Prospect Theory

## Probability Weighting

```{r echo=FALSE, message=FALSE, warning=FALSE}
cols <- list(.default = col_double(),
             boundary = col_factor(),
             a = col_factor(),
             parameter = col_factor())
cpt <- read_csv("data/estimates/estimates_cpt.csv", col_types = cols)

cpt <- cpt %>% 
  select(s, boundary, a, parameter, mean) %>% 
  pivot_wider(names_from = parameter, values_from = mean)

cpt_w <- cpt %>% 
  select(-c(alpha, rho)) %>% 
  expand_grid(ep = seq(0, 1, .1)) %>% #
  mutate(w = round(  (delta * ep^gamma)/ ((delta * ep^gamma)+(1-ep)^gamma), 2))

## plot curves
curve_w <- cpt_w %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path() + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(title = "Probability Weighting", 
       x ="Experienced Probability p",
       y = "w(p)", 
       color="Switching Probability") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom")

gamma <- cpt %>% 
  group_by(boundary, a) %>% 
  ggplot(aes(x = s, y = gamma, color = s)) +
  geom_smooth(data = cpt %>% filter(a != 1), color = "grey") + 
  geom_jitter(size = 3, show.legend = FALSE) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  labs(title = expression(gamma),
       x ="",
       y = "") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

delta <- cpt %>% 
  ggplot(aes(x = s, y = delta, color = s)) +
  geom_smooth(data = cpt %>% filter(a != 1), color = "grey") + 
  geom_jitter(size = 3, show.legend = FALSE) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(limits = c(1, 2), breaks = seq(1, 2, .2)) +
  labs(title = expression(delta),
       x ="Switching Probability",
       y = "") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

params <- ggarrange(gamma, delta, nrow = 2)
ggarrange(curve_w, params)
```

## Probability Weighting: Multiple Comparisons

```{r echo=FALSE}
cpt_w %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path(size = 1.2) + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(x ="Experienced Probability p",
       y = "w(p)", 
       color="Switching Probability") + 
  facet_wrap(~a, nrow = 2) +  
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom")
```

## False Response Rates

```{r echo=FALSE, message=FALSE, warning=FALSE}
cols <- list(.default = col_double(),
             boundary = col_factor(),
             gamble = col_factor(),
             rare = col_factor(),
             agent = col_factor(),
             choice = col_factor())
choices <- read_csv("data/choices/choices.csv", col_types = cols)

# compute false response rates

fr_rates <- choices %>% 
  mutate(ev_ratio_exp = round(a_ev_exp/b_ev_exp, 2), # experienced EV (eEV)
         norm = case_when(ev_ratio_exp > 1 ~ "A", ev_ratio_exp < 1 ~ "B")) %>% # normative choice according to eEV
  filter(!is.na(norm)) %>% # exclude trials with normative indifferent prospects
  group_by(s, boundary, a, rare, norm, choice) %>% # group correct and incorrect responses
  summarise(n = n()) %>% # absolute numbers 
  mutate(rate = round(n/sum(n), 2), # response rates 
         type = case_when(norm == "A" & choice == "B" ~ "false safe", norm == "B" & choice == "A" ~ "false risky")) %>% filter(!is.na(type))  # remove correct responses

fr_rates %>% 
  ggplot(aes(s, a, fill = rate)) + 
  geom_tile(colour="white", size=0.1) +
  scale_fill_scico(palette = "buda") + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  scale_x_continuous(expand=c(0,0), breaks = seq(0, 1, .1)) +
  scale_y_continuous(expand=c(0,0), breaks = seq(1, 5, 1)) +
  labs(title = "Type of Rare Event", 
       x = "Switching Probability", 
       y= "Number of Comparisons", 
       fill = "% False Responses") + 
  theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
```

## Value Function

```{r echo=FALSE, message=FALSE, warning=FALSE}
cpt_v <- cpt %>% 
  select(-c(gamma, delta, rho)) %>% 
  expand_grid(x = seq(0, 20, 2)) %>%  
  mutate(v = round(x^alpha, 2))

## plot curves
curve_v <- cpt_v %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = x, y = v)) +
  geom_path(aes(color = s)) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(breaks = seq(0, 20, 2)) + 
  labs(title = "Value Function",
       x ="Objective Value x",
       y = "v(x)", 
       color="Switching Probability") + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom")

alpha <- cpt %>% 
  group_by(boundary, a) %>% 
  ggplot(aes(x = s, y = alpha, color = s)) +
  geom_smooth(data = cpt %>% filter(a != 1), color = "grey") + 
  geom_jitter(size = 3, show.legend = FALSE) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(breaks = seq(0, 2, .1)) +
  labs(title = expression(alpha),
       x ="Switching Probability",
       y = "") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(curve_v, alpha)
```

## Value Function: Multiple Comparisons

```{r echo=FALSE, message=FALSE, warning=FALSE}
cpt_v %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = x, y = v, color = s)) +
  geom_path(size = 1.2) + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(breaks = seq(0, 20, 2)) + 
  labs(x ="Objective Value x",
       y = "v(x)", 
       color="Switching Probability") + 
  facet_wrap(~a, nrow = 2) +  
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom")
```

## Summary and Q&A

Random processes underlying DfE ...

- ... can be modeled according to probability theory

- ... can be integrated into current decision theories

- ... could be shaped by sampling and decision strategies

- ... could explain empirically observed choice patterns

I happily take your questions, comments and critique.

## References