### 2022-03-11 manuscript_appendix.Rmd
Add appendix with anlyses of comprehensive simulation

```{r A-setup, include=FALSE}
knitr::opts_chunk$set(echo = F, #omit messages
                      fig.align = "l", #align figures to left margin
                      fig.pos = "t")
```

\begin{figure}
\caption{Effect of Switching Probability on Rates of False Risky and False Safe Choices for 2-Outcome Prospects with High Rank Outcomes of Different Probability}\label{fig:A-choice-rates}
\includegraphics{manuscript_files/figure-latex/A-choice-rates-1}
\textit{Note.} A note.
\end{figure}

\newpage

\begin{figure}
\caption{Parameter Estimates of the Weighting Function of Goldstein and Einhorn (1987) for Different Simulation Parameters}\label{fig:A-weighting-parameters}
\includegraphics{manuscript_files/figure-latex/A-weighting-parameters-1}
\end{figure}

\newpage

\begin{figure}
\caption{Parameter Estimates of the Value Function of Tversky and Kahneman (1992) and of the Logit Choice Rule for Different Simulation Parameters}\label{fig:A-value-parameters}
\includegraphics{manuscript_files/figure-latex/A-value-parameters-1}
\end{figure}

\newpage

\begin{figure}
\caption{Estimated Graphical Shapes of the Weighting and Value Function for Different Simulation Parameters}\label{fig:A-cpt-graphs}
\includegraphics{manuscript_files/figure-latex/A-cpt-graphs-1}
\end{figure}